### 2022-11-02 Tutorial_code.Rmd
Tutorial code

---
title: "Untitled"
output: html_document
date: "2022-11-02"
---

```{r}
#setting up variables
n=1000
x=rnorm(n,0,1)
p=exp(0.1+0.5*x)/(1+exp(0.1+0.5*x))
y=rbinom(n,1,p)
con=rep(1,n)
X=matrix(c(con,x),nrow=n,ncol=2)

# #Fiddling around to work out the matrix algebra
# T <- matrix(c(1,2,3,4),ncol=2,nrow=2,byrow=T)
# beta <- matrix(c(1,1),ncol=1,nrow=2)
# beta
# diag(as.vector(exp(T%*%beta)/(1+exp(T%*%beta))^2),2)
#   
# vec <- c(1,0)-as.vector(exp(T%*%beta)/(1+exp(T%*%beta)))
# t(vec%*%T)
```

```{r}
# First make with iteration to be specified to check function is working
MLE.iter <- function(y,X,iter){

  beta.previous <- matrix(c(1,1),nrow=2,ncol=1)

  for (i in 1:iter){
    beta <- beta.previous
    W <- diag(as.vector(exp(X%*%beta)/(1+exp(X%*%beta))^2),dim(X)[1])
    Y <- y-as.vector(exp(X%*%beta)/(1+exp(X%*%beta)))

    beta.previous<- beta+solve(t(X)%*%W%*%X)%*%t(Y%*%X)
  }
  return(beta.previous)
}

#Using convergence this time to stop iterations (or if reaches max iterations)
MLE.conv <- function(y,X){
  #Setting initial guess of beta
  beta.new <- matrix(c(1,1),nrow=2,ncol=1)
  #Setting objects for conditions of when stop iterations (while loop)
  flag <- 1
  eps <- 0.00001
  c <- 0
  max.iter <- 100

  while (flag==1){
    #Setting beta to use to find new value of beta
    beta <- beta.new

    #Finding W, where I have defined W as t(X)WX = Fisher information matrix
    W <- diag(as.vector(exp(X%*%beta)/(1+exp(X%*%beta))^2),dim(X)[1])
    #Finding Y where t(YX) is defined as the score 
    Y <- y-as.vector(exp(X%*%beta)/(1+exp(X%*%beta)))

    #Applying Fisher scoring
    beta.new<- beta+solve(t(X)%*%W%*%X)%*%t(Y%*%X)

    #Checking for convergence
    if (abs(sum(beta.new)-sum(beta))<eps){
      flag <- 0
    }
    c <- c+1
  }

  var.cov.mat <- solve(t(X)%*%W%*%X)

  return(list("estimates"=beta.new,"vcov"=var.cov.mat))
}

MLE.output <- MLE.conv(y,X)
MLE.output$estimates

#Testing against glm function
glm.model <- glm(y~x,family = "binomial")
summary(glm.model)

#Checking variance-covariance matrix against that found in glm function
vcov(glm.model)
MLE.conv(y,X)$vcov
```