### 2022-11-04 Task_2_draft.Rmd
adding comments in code

---
title: "Untitled"
output: html_document
date: "2022-11-02"
---
```{r}
#Setting up enviroment
setwd("~/Desktop/GLM/Coursework/GLM-CW")
airline.df <- read.table("airline.txt",header=T)

#Setting up inputs of mle function
n <- dim(airline.df)[[1]]
con <- rep(1,n)
X=matrix(c(con,airline.df$x),nrow=n,ncol=2)
y <- airline.df$y
```

```{r}
#Creating function for algorithm
Fisher.scoring.alg <- function(X,y){
  #Setting initial guess of beta coefficients

  beta.new <- matrix(c(0,1),nrow=2,ncol=1)
  #Setting objects for conditions of when stop iterations (while loop)
  flag <- 1
  #setting size of absolute difference between beta iterations when convergence is reached
  eps <- 0.0000001
  #Setting loop counter
  c <- 0
  #Max iterations before 
  max.iter <- 1000000

  while (flag==1){
    #Updating current value of beta to use for current iteration to find next beta value
    beta <- beta.new

    #Updating Y where YX is the score function
    Y <- as.vector((y*exp(-(X%*%beta)))-1)

    #Applying Fisher Scoring algorithm to find updated value of beta 
    beta.new <- beta + solve(t(X)%*%X)%*%t(Y%*%X)

    #Checking for convergence
    if (abs(sum(beta.new)-sum(beta))<eps){
      flag <- 0
    }

    if (c>max.iter){
      stop("Max number of iterations reached: Non-convergence of beta")
    }
    c <- c+1
  }
  #Variance covariance matrix
  var.cov.mat <- solve(t(X)%*%X)

  return(beta.new)
}

Fisher.scoring.alg(X,y)
```

```{r}
#Check with glm model
fit <- glm(y~X[,2], family  = Gamma(link="log"))
summary(fit,dispersion=1)  
```

```{r}
#Finding var-cov matrix

t(X)%*%(X)
```