### 2022-11-06 Draft_1.Rmd
Large changes to model selection process (now fairly happy with final model chosen)

---
title: "Untitled"
output:
  pdf_document: default
  html_document: default
date: "2022-10-27"
---
```{r,echo=F}
#Initial environment setup
setwd("~/Desktop/GLM/Coursework/GLM-CW")
expend.df <- read.table("expenditure.txt")
library(tidyverse)
library(kableExtra)
library(gridExtra)
library(broom)
library(MASS)
```

```{r,echo=F}
#Data cleaning

#coding variables as factors
expend.df$house.ten <- factor(expend.df$house.ten,levels=c(1:3),labels=c("Public Rented","Private rented","Owned"))
expend.df$sex.hh <- factor(expend.df$sex.hh,levels=c(1,2),labels=c("Male","Female"))
expend.df$lab.force <- factor(expend.df$lab.force,levels=c(1:4),labels=c("Full time", "Part time","Unemployed","Inactive"))
expend.df$hh.size <- factor(expend.df$hh.size,levels=c(1:5),labels = c("1 person","2 people","3 people","4 people","5+ persons"))
expend.df$hh.adults <- factor(expend.df$hh.adults,levels = c(1:4), labels = c("1 adult","2 adults","3 adults","4+ adults"))

```
## Task 1:

#### Part 1:
```{r,echo=F}
#Histogram to examine the expenditure
hist(expend.df$expenditure,breaks = 20,
     xlab = "Total weekly household expenditure (GBP)",
     main = "Histogram of expenditure")

#Creating the 
#scatter plot to look at relationship between income and expenditure
p1<-ggplot(data=expend.df,aes(x=income,y=expenditure))+
  geom_point(size=0.2)+
  xlab("Gross weekly average household income (GBP)")+
  ylab("Total expenditure")+
  theme(axis.text=element_text(size=6),
        axis.title=element_text(size=6,face="bold"))

#boxplots
p2<-ggplot(data=expend.df,aes(y=expenditure,x=house.ten,fill=house.ten))+
  geom_boxplot()+
  xlab("Household tenure")+
  guides(fill=F)

#scale_fill_discrete(name="Household\ntenure")

p3<-ggplot(data=expend.df,aes(y=expenditure,x=sex.hh,fill=sex.hh))+
  geom_boxplot()+
  xlab("Sex of the household head")+
  guides(fill=F)

p4<-ggplot(data=expend.df,aes(y=expenditure,x=lab.force,fill=lab.force))+
  geom_boxplot()+
  xlab("Employment status")+
  guides(fill=F)

p5<-ggplot(data=expend.df,aes(y=expenditure,x=hh.size,fill=hh.size))+
  geom_boxplot()+
  xlab("Household size")+
  guides(fill=F)

p6<-ggplot(data=expend.df,aes(y=expenditure,x=hh.adults,fill=hh.adults))+
  geom_boxplot()+
  xlab("Number of adults in the household")+
  guides(fill=F)

grid.arrange(p1,p2,p3,p4,p5,p6,ncol=2,nrow=3)
```

Looking at the above histogram of expenditure above it shows that expenditure has a positively skewed distribution.
Investigating the relationship between household income and expenditure in the scatter plot above, it displays a clear positive association.
Then to investigate the other variables which are categorical I used boxplots to display differences in the expenditure across the different levels(categories). The differences shown in the plots are listed below:

-Different household tenures categories have different weekly expenditure amounts with public rented households having a noticeably lower expenditure compared to the other 2 household levels.
-Households with a female head have a lower weekly expenditure compared to male household heads
-Employment status appears to be associated with expenditure with unemployed households having a dramatically lower expenditure compared to households with full or part time employment.
-Household size generally appears to be positively correlated with expenditure, with larger households having a higher weekly expenditure (however 5+ households break this trend with the expenditure decreasing, this could be due to other variables).
Number of adults in the household also appears positively correlated with expenditure with households containing more adults having a higher expenditure.

I note that these differences may not be due to the variable mentioned but due to a third variable (a confounder) that affects both the categorical variable and expenditure (the response) which could be other variables in the dataset or ones that were not even included.

#### Part 2:
In all linear models mentioned for part 2 to 6, the response $y$ is the weekly household expenditure and the explanatory variable $x$ is weekly income and $i \in \{1,...,1200\}$ is the ith household in the dataset.

```{r,echo=F}
#Question 2
lm.income<-lm(expenditure~income,data = expend.df)
#summary(lm.income)
as.data.frame(tidy(lm.income)[,1:3])%>%
  kable(escape = F)%>%
  kable_styling(font_size = 15)
par(mfrow=c(2,2))
plot(lm.income)
```
The estimated coefficients for the fitted linear model $y_{i}=\beta_{0}+\beta_{1}x_{i}+\epsilon$  are displayed in the above table. It shows the estimated values of the intercept ($\beta_0$) and the estimated 'effect' of a 1 pound increase in income on expenditure $\beta_1$ and the associated standard errors.

Viewing the diagnostic plots for this model. The fitted vs residual plot shows the residuals are not equally spread around 0 with a greater spread of residuals above zero, meaning the assumption of linearity may not hold. Furthermore, the QQ normal plot has a u shape with the tails diverging heavily away from the straight line. This implies the normality assumption is invalid.

#### Part 3:
```{r,echo=F}
#Question 3
lm.income.sqd <- lm(expenditure~income+I(income^2),data = expend.df)
tidy(lm.income.sqd)[,1:3]%>%
  kable(escape = F)%>%
  kable_styling(font_size = 15)
par(mfrow=c(2,2))
plot(lm.income.sqd)

# lm.log.income.sqd <- lm(log(expenditure)~income+I(income^2),data = expend.df)
# summary(lm.log.income.sqd)
# plot(lm.log.income.sqd)
# 
# gamma.income.sqd <-glm(expenditure~income+I(income^2),data=expend.df,family="Gamma")
# summary(gamma.income.sqd)
# plot(gamma.income.sqd)
```
Fitting a new linear model $y_{i}=\beta_{0}+\beta_{1}x_{i}+\beta_2x^2_{i}+\epsilon$ with an added squared term of income gives the above estimated coefficients. Looking at the diagnostic plots, in the fitted vs residual plot the residual values increase as the fitted values increase thus the homoscedasticity assumption may not be valid. The tails in the normal QQ plot also still deviate heavily away from the straight line meaning the normality assumption is invalid.

#### Part 4:
```{r,echo=F}
#Question 4
lm.log.income<-lm(log(expenditure)~income,data = expend.df)
tidy(lm.log.income)[,1:3]%>%
  kable(escape = F)%>%
  kable_styling(font_size = 15)
par(mfrow=c(2,2))
plot(lm.log.income)
# #QQ plot much better but not sure about the residual vs fitted if the variance decreases as fitted value increases
```
Fitting the same model as in task 2 but with the response of expenditure logged $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\epsilon$ gives the above coefficients. The data appears to be heteroscedastic as the fitted vs residual plot points spread decreases as fitted values increases, furthermore the linear assumption also appears violated as the residuals appear to exhibit a slight negative quadratic shape. The QQ plot shows the residuals stay close to the straight line thus normality assumption is valid.

#### Part 5:
```{r,echo=F}
lm.log.income.sqd <- lm(log(expenditure)~income+I(income^2),data = expend.df)
tidy(lm.log.income.sqd)[,1:3]%>%
  kable(escape = F)%>%
  kable_styling(font_size = 15)
par(mfrow=c(2,2))
plot(lm.log.income.sqd)
```
Fitting the model $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\beta_{2}x^2_{i}+\epsilon$ gives the estimated coefficients above. Viewing the diagnostic plots the fitted vs residuals plot shows a fairly even scattering of points around 0 and a linear trend, however the residuals are bunched to the very right of the plot and the variance of the residuals appears to decrease as the fitted values increase, this could be to important variables being left out of the model (MAY NEED TO CHANGE THIS). The non-constant variance assumption being violated is further backed up by the none straight line in the scale location plot. The QQ plot shows the points stay close to the straight line meaning the normality assumption is valid.

#### Task 6
```{r}
#Comparing AIC
paste("AIC")
AIC(lm.income)
AIC(lm.income.sqd)
AIC(lm.log.income)
AIC(lm.log.income.sqd)

# #Comparing R squared
# glance(lm.income)$r.squared
# glance(lm.income.sqd)$r.squared
# glance(lm.log.income)$r.squared
# glance(lm.log.income.sqd)$r.squared

#Comparing adjusted R squared
paste("adj R squared")
glance(lm.income)$adj.r.squared
glance(lm.income.sqd)$adj.r.squared
glance(lm.log.income)$adj.r.squared
glance(lm.log.income.sqd)$adj.r.squared
```

The 4th model $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\beta_{2}x^2_{i}+\epsilon$ has the lowest AIC value of *1528.062* and the highest adjusted R squared value of *0.4635055*. Furthermore, looking at the diagnostic plots the normality assumption is valid and the fitted vs residual plot although not 'perfect' displays a more even scattering of points than that of the same plot for the other models. Therefore for all these reasons the 4th model is my preferred model to describe the relationship between expenditure and income.

```{r}
tidy(lm.log.income.sqd)

mean(expend.df$income)
```
Income has a positive association with expenditure with a 1 pound increase in household income increasing household expenditure by $exp(0.0032-0.0000029x-0.0000015)-1$x100%. Therefore for a household earning the mean income of £512 a £1 increase in income increases expenditure by approximately 0.17%. Using a more likely realistic increase in weekly income of £20 for a household earning £512 gives an increase of $exp(0.016-0.000015*20-0.000036)-1$x100% = 1.6%.
#### Task 7
To start I investigated the multicolinearity of the variables to make sure that certain variables were not linearly dependent as this would cause issues with the model if linearly dependent variables were included.

```{r}
#Checking variance inflation factors (generalized as variables include factors)
car::vif(lm(log(expenditure)~income+I(income^2)+hh.size+hh.adults+sex.hh+lab.force+house.ten,data=expend.df))
```
None of the variables had a GVIF greater than 5, thus multicolinearity is not an issue with the explanatory variables. Thus multicolinearity will not be an issue. However, considering the variables it seems likely that household size is related to number of household adults.

```{r}
#Creating tables for chi-squared test
prop.table(table(expend.df$hh.size,expend.df$hh.adults),2)
```
It is clear there is a relationship between the two. I now create linear models includiong each seperately and both together.

```{r}
lm.adult <- lm(log(expenditure)~income+I(income^2)+hh.adults,data = expend.df)
summary(lm.adult)
#adult significant

lm.size <- lm(log(expenditure)~income+I(income^2)+hh.size,data = expend.df)
summary(lm.size)
#size significant

lm.adult.size <- lm(log(expenditure)~income+I(income^2)+hh.size+hh.adults,data = expend.df)
summary(lm.adult.size)
#adult no longer significant
```

```{r}
# #Checking which of the correlated variables to include
# lm.hh.size <- lm(log(expenditure)~income+I(income^2)+hh.size, data = expend.df)
# lm.hh.adults <- lm(log(expenditure)~income+I(income^2)+hh.adults, data = expend.df)
# 
# AIC(lm.hh.size)
# AIC(lm.hh.adults)
# #size has lower AIC
# 
# glance(lm.hh.size)$r.squared
# glance(lm.hh.adults)$r.squared
# #Size has higher R^2

#Based on AIC and R^2 (can use as all models have same number of variables) I select household size and tenure as the 2 variables to use in the model and won't use labour or household adults variables so will no longer consider these 2 variables.
```
Looking at the summary outputs of all 3 variables the number of adults in the household is a significant predictor when on its own for predicting expenditure however when we add size of household it is no longer significant. Therefore the effect of number of adults can be explained by the size fo the household thus there is no point including it in the model. Thus, I will not consider the number of adults in the household as an explanatory variable for the model.

Considering the possible interactions, I decided to only include 2 way interactions as higher order interactions make interpretation of the model much more complicated and there is no clear 3 way interactions that make sense.

First I investigated interactions between income and the factor variables. To do this I created scatter plots of income against expenditure with income split into each factor of the chosen explanatory factor variable with a plotted linear model for each of the level of the factor to see if the effect of income varied with each level of the factor variable on expenditure. This would be seen by a different gradient of the linear model.

```{r}
#interaction between 3 factor variables and income investigation
interaction.continous <-function(var){
  ggplot(data=expend.df,aes(x=income,y=expenditure,colour=.data[[var]]))+
    geom_point()+
  geom_smooth(method='lm', formula= y~x,se=F)
}

int1<-interaction.continous("house.ten")
int2<-interaction.continous("sex.hh")
int3<-interaction.continous("hh.size")
int4<-interaction.continous("lab.force")

grid.arrange(int1,int2,int3,int4,ncol=2,nrow=2)
int1
int2
int3
int4
```
Looking at the plots all show that factor variables have an affect on the effect of income on expenditure with different gradients between the levels, however the magnitude of the difference in these gradients varies. With sex the difference in gradient is very small indicating that sex has little affect on the effect of income on expenditure. Labour force on the hand has a large difference in gradients between its levels giving evidence of an interaction between labour force and income. I decided to not include a interaction between sex and income as there is little evidence to support this (and makes logical sense).

Looking at the plots. The largest differences in gradients between levels are in income:labour force, income:household adults and income: household size plots, thus these are the most likely to have interactions with income. Furthermore, it makes logical sense that these variables would affect to effect of income on expenditure. Meanwhile sex has little effect on income with the gradients although crossing remaining very similar and house tenure type having some differences.

```{r}
#Checking each of the interactions to see if significant with all variables included
# lm.int.income.sex <- lm(log(expenditure)~income+I(income^2)+hh.size+sex.hh*income+lab.force+house.ten,data=expend.df)
# summary(lm.int.income.sex)
# #not significant as expected

lm.int.income.size <- lm(log(expenditure)~income+I(income^2)+hh.size*income+sex.hh+lab.force+house.ten,data=expend.df)
summary(lm.int.income.size)
#significant

#significant
lm.int.income.labour <- lm(log(expenditure)~income+I(income^2)+hh.size+sex.hh+lab.force*income+house.ten,data=expend.df)
summary(lm.int.income.labour)
#significant

lm.int.income.tenure <- lm(log(expenditure)~income+I(income^2)+hh.size+sex.hh+lab.force+house.ten*income,data=expend.df)
summary(lm.int.income.tenure)
#significant
```
Considering factor variable interactions. There are 6 possible factor interactions.

```{r}
# attach(expend.df)
# prop.table(table(sex.hh,hh.size),1)
# prop.table(table(sex.hh,hh.size),2)
# #differences
# 
# prop.table(table(sex.hh,house.ten),1)
# prop.table(table(sex.hh,house.ten),2)
# 
# prop.table(table(sex.hh,lab.force),1)
# prop.table(table(sex.hh,lab.force),2)
# #differences
# 
# prop.table(table(house.ten,hh.size),1)
# prop.table(table(house.ten,hh.size),2)
# 
# prop.table(table(lab.force,hh.size),1)
# prop.table(table(lab.force,hh.size),2)
# 
# prop.table(table(lab.force,house.ten),1)
# prop.table(table(lab.force,house.ten),2)
```

```{r}
lm.int.sex.size <- lm(log(expenditure)~income+I(income^2)+hh.size*sex.hh+lab.force+house.ten,data=expend.df)
summary(lm.int.sex.size)
#significant

lm.int.sex.labour <- lm(log(expenditure)~income+I(income^2)+hh.size+sex.hh*lab.force+house.ten,data=expend.df)
summary(lm.int.sex.labour)
#not significant

lm.int.sex.tenure <- lm(log(expenditure)~income+I(income^2)+hh.size+sex.hh*house.ten+lab.force,data=expend.df)
summary(lm.int.sex.tenure)
#not significant

lm.int.size.labour <- lm(log(expenditure)~income+I(income^2)+hh.size*lab.force+sex.hh+house.ten,data=expend.df)
summary(lm.int.size.labour)
#not significant

lm.int.size.tenure <- lm(log(expenditure)~income+I(income^2)+hh.size*house.ten+sex.hh+lab.force,data=expend.df)
summary(lm.int.size.tenure)
#not significant

lm.int.labour.tenure <- lm(log(expenditure)~income+I(income^2)+hh.size+sex.hh+lab.force*house.ten,data=expend.df)
summary(lm.int.labour.tenure)
#significant
```
Adding each to the basic model separately only sex:household size and labour:tenure were significant therefore these are the only 2 I will consider for the final model.

#### From full model
To choose my final model I will start with a full model that includes all possible variables that I want to consider. Then using AIC and adjusted R^2 I will remove the coefficients. (lower AIC or higher adjusted R^2). Once I am happy with the model I end up with I will check the diagnostic plots to make sure the model fit is good/ assumptions met.

```{r}
lm.log.full<-lm(log(expenditure)~income*lab.force+I(income^2)+income*house.ten+income*hh.size+sex.hh:hh.size+lab.force:house.ten,data=expend.df)
step.method<- step(lm.log.full,trace=F)
tidy(step.method)
#labour force and household tenure interaction removed
summary(step.method)

#comparing with original model
AIC(lm.log.income.sqd)
AIC(step.method)

glance(lm.log.income.sqd)$adj.r.squared
glance(step.method)$adj.r.squared

anova(step.method)
#All variables left statistically significant
```
Using AIC as the model selection criteria I removed 1 interactions labour force:house tenure type. Comparing this new model to the original the AIC is significantly lower and the adjusted R squared is significantly higher. All variables left in the model are also statistically significant.

Creating diagnostic plots
```{r}
par(mfrow=c(2,2))
plot(step.method)
```
All look good... EXPLAIN

```{r}
# #Comparing to no interaction model
# lm.log.full.no.int<-lm(log(expenditure)~income+I(income^2)+sex.hh+house.ten+lab.force+hh.size,data=expend.df)
# summary(lm.log.full.no.int)
# step.method<- step(lm.log.full.no.int,trace=F)
# 
# plot(lm.log.full.no.int)
# AIC(lm.log.full.no.int)
# #higher AIC

# glance(lm.log.full.no.int)$adj.r.squared
# #lower adjusted R^2
```

## Task 8