### 2022-11-18 Deleted_stuff.Rmd
adding chnages

---
title: "Untitled"
output: html_document
date: "2022-10-29"
---
```{r,echo=F,eval=F}
#extra code that may not use
pairs(expend.df)
xtabs(~expend.df$house.ten+expend.df$sex.hh)

par(mfrow=c(1,2))
hist(expend.df$expenditure,breaks = 20,
     xlab = "Total weekly household expenditure (GBP)")
hist(log(expend.df$expenditure),breaks = 20,
     xlab = " Log of total weekly \n household expenditure (GBP)")
#Applying log transform gives a more normal looking distribution
```

```{r}
# #as expected the normal QQ plot shows 'bad fit'

# #try gamma as good for postively skewed continous data
# gamma.income <-glm(expenditure~income,data=expend.df,family="Gamma")
# summary(gamma.income)
# plot(gamma.income)
# #bad fit (explain)
# 
# #out of all models log linear is the most appropriate
```

```{r}
#Creating model with all possible 2 way interactions (not a good idea but just for interest :P)
lm.log.full<-lm(log(expenditure)~income+I(income^2)+income*sex.hh+income*hh.adults+income*house.ten+income*lab.force+income*hh.size+sex.hh*hh.adults+sex.hh*house.ten+sex.hh*lab.force+sex.hh*hh.size+hh.adults*house.ten+hh.adults*lab.force+hh.adults*hh.size+house.ten*lab.force+house.ten*hh.size+lab.force*hh.size,data=expend.df)
summary(lm.log.full)

step.method<- step(lm.log.full,trace=F)
summary(step.method)

```